\documentclass{article}

%some useful packages
\usepackage{amsmath,amsthm,amsfonts,amssymb}  %math stuff
\usepackage{graphicx} %to embed images
\usepackage{enumitem} %fancy lists
\usepackage{fullpage} %smaller margins
\usepackage{color} %smaller margins

%some common macros.
\renewcommand{\P}{\mathbb P}
\newcommand{\Z}{\mathbb Z}
\newcommand{\E}{\mathbb E}
\newcommand{\R}{\mathbb R}
\newcommand{\N}{\mathbb N}
\newcommand{\eps}{\varepsilon}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\setlist[itemize,1]{nosep}
\setlist[enumerate,1]{nosep,label=(\alph*)}

\newenvironment{solution}
{\paragraph{\color{blue}Solution.}}{\medbreak}

\usepackage{fancyhdr}
\fancypagestyle{crfooter}{%
  \fancyhf{}
  \renewcommand\headrulewidth{0pt}
  \fancyfoot[R]{\copyright\ Omer Angel 2021, all rights reserved.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\thispagestyle{crfooter}

\begin{center}
  {\LARGE \textbf{Mathematical Game theory}} \\
  {\Large Assignment 1, due 2022-01-21}
\end{center}

\hrule width \textwidth \bigskip

\begin{problem}
  A Markov chain with state space $S=\{0,1,2\}$ has the transition matrix
  \[
    P =
    \begin{pmatrix}
      .2 & .3 & .5 \\
      0 & .4 & .6 \\
      .7 & .3 & 0
    \end{pmatrix}.
  \]
  If $X_0$ has distribution vector $(.6,.4,0)$, find the distribution vectors of $X_1,X_2$.
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
  There are 6 coins on a table, each showing heads (H) or tails (T).
  In each step we 
  \begin{itemize}
  \item Select uniformly one of the coins.
  \item If it is heads, toss it and replace on the table (with random side).
  \item If it is tails, toss it.
    If it comes up heads leave it at that.
    If if comes up tails, toss it a second time, and leave the result as it is.
  \end{itemize}
  Let $X_n$ be the number of heads showing after $n$ such steps.
  \begin{enumerate}
  \item Determine the transition probabilities for this Markov chain.
  \item Draw the transition diagram and write the transition matrix.
  \item What is $\P(X_2=4 | X_0=5)$?
  \end{enumerate}
\end{problem}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}
  Jaine is a mechanic.
  Let $X_n$ be the number of cars they need to fix.
  At each step, with probability 1/2 they manage to fix a car, and $X$ decreases by 1.
  With probability 1/3 a new car is delivered for service and $X$ increases by 1, and with probability 1/6 nothing happens and $X$ stays the same.
  If $X$ becomes 0, Jaine goes for lunch.
  If $X$ becomes 3, the boss gets angry and fires Jaine.

  \begin{enumerate}
  \item Suppose $X_0=1$.
    What is the probability that Jaine goes to lunch before they are fired?
  \item What would be the answer if the boss only fires Jaine when $X=4$?
  \end{enumerate}
\end{problem}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}
  Consider the simple random walk on a triangle with vertices $\{0,1,2\}$.
  This is the Markov chain that at each step moves to one of the two states other than $X_n$.
  Prove that the $n$-step transition matrix is
  \[
    P^n = 
    \begin{pmatrix}
      a_n & b_n & b_n \\
      b_n & a_n & b_n \\
      b_n & b_n & a_n 
    \end{pmatrix}, \]
  where $a_n = \frac{1+2(-1/2)^n}{3}$ and $b_n = \frac{1-(-1/2)^n}{3}$.
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}
  Alfonso, Beatrice, and Coraline play poker.
  They start with \$1, \$2, \$3 respectively.
  In each round, one of the three (chosen randomly) does not participate.
  The other two each bet \$1, and one of the two wins the bet (and the other loses), again randomly and equally likely.
  If a player has no money, they are eliminated and do not participate in future rounds.
  Eventually one player has all \$6, and the tournament ends.
  \begin{enumerate}
  \item What is the probability that Alfonso is the eventual winner?
    (Hint: compare Alfonso's money to the gambler's ruin problem.)
  \item Let $q_A$ be the probability that Alfonso is the first one to be eliminated, and similarly $q_B,q_C$ for Beatrice and Coraline.
    Compute $q_A,q_B,q_C$.
    (Hint: Consider also what happens if the players have \$1,\$1,\$4.)
  \end{enumerate}
\end{problem}

\paragraph{Extra practice problems}
Do not hand these in. (Feel free to ask for hints is stuck.)

Ross, chapter 4: problems 1,3,4,6,8,9.

\end{document}
