\documentclass{article}

%some useful packages
\usepackage{amsmath,amsthm,amsfonts,amssymb}  %math stuff
\usepackage{graphicx} %to embed images
\usepackage{enumitem} %fancy lists
\usepackage{fullpage} %smaller margins
\usepackage{color} %smaller margins

%some common macros.
\renewcommand{\P}{\mathbb P}
\newcommand{\Z}{\mathbb Z}
\newcommand{\E}{\mathbb E}
\newcommand{\R}{\mathbb R}
\newcommand{\N}{\mathbb N}
\newcommand{\eps}{\varepsilon}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Poi}{Poi}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Exp}{Exp}
\DeclareMathOperator{\Bin}{Bin}
\DeclareMathOperator{\Geom}{Geom}
\DeclareMathOperator{\Bernoulli}{Bernoulli}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\setlist[itemize,1]{nosep}
\setlist[enumerate,1]{nosep,label=(\alph*)}

\newenvironment{solution}
{\paragraph{\color{blue}Solution.}}{\medbreak}

\usepackage{fancyhdr}
\fancypagestyle{crfooter}{%
  \fancyhf{}
  \renewcommand\headrulewidth{0pt}
  \fancyfoot[R]{\copyright\ Omer Angel 2022, all rights reserved.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\thispagestyle{crfooter}

\begin{center}
  {\LARGE \textbf{Stochastic Processes}} \\ \medskip
  {\Large Assignment 6, due 2022-03-11}
\end{center}

\hrule width \textwidth

\paragraph{Note:} Start each problem on a \textbf{new page}.% (Multiple pages for a single problem are fine if needed.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
  In a pizzeria the orders arrive according to a Poisson process of rate $\lambda$. Given that 3 orders arrived in an hour, determine the conditional probability density functions of the arrival of the first times $S_1,S_2,S_3$.
  (Note: this asks for 3 probability density functions, not the joint p.d.f.)
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
  Let $N(t)$ be a Poisson process of rate $\lambda$.
  Determine the conditional distribution of $N(3)$ given that $N(5)=7$.
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
  Raine is a birdwatcher and counts the birds in her garden, which arrive according to a Poisson process of rate $\lambda$.
  If an hour passes with no birds, she gets bored and goes to make tea.
  \begin{enumerate}
  \item What is the probability that she doesn't see any birds?
  \item How much time on average does she spend birdwatching?
    Hint: Condition on the time of the first observed bird.
  \item What is the distribution of the number of birds seen?
  \end{enumerate}
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
  Players arrive at a casino at rate $\lambda$, and begin betting, and losing money at rate \$1/minute.
  Let $X$ be the total amount they lose by time $t$.
  \begin{enumerate}
    \item What is $E(X|N(t))$?
    \item Explain why $\Var(X|N(t)) = N(t) t^2/12$.
    \item Find $\Var(X)$.
  \end{enumerate}
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
  A system suffers faults by a Poisson process with rate $\lambda$.
  Each fault will, independently, cause the system to fail after time $X$ with
  c.d.f. $F$.
  When the system fails, the responsible fault is fixed, but earlier other faults may remain and cause later failure.
  \begin{enumerate}
  \item What is the distribution of the number of failures up to $t$?
  \item What is the distribution of the number of faults that remain in the system at time t and are not yet fixed?
  \item Are the random variables in parts (a) and (b) dependent or independent? (justify!)
  \end{enumerate}
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hrule width \textwidth

\paragraph{Extra practice problems}
Do not hand these in. (Feel free to ask for hints is stuck.)

Ross, chapter 5: problems 50, 53, 56, 64,68, 71.

\paragraph{Read ahead}
We will start Chapter 6 next week.

\end{document}
